\subsection{The basic definitions and conventions}

\noindent The terms \highlight{adjoint} and \highlight{dual} are often used interchangeably.
We will standardize on \highlight{adjoint}, to avoid unnecessary confusion.
Following\ \cite{werner_funktionalanalysis_2018}, we write $\thefield \in \{\reals, \complexes\}$ when the field is unspecified.

\newcommand{\twonormedvectorspaces}{$\normedspace{X}$ and $\normedspace{Y}$ be normed \thefield-vector spaces}
\begin{definition}
    We remind outselves of the following concepts from the lecture \cite{krieg_functional_2025}:
    \begin{romanenum}
        \item
            Let \twonormedvectorspaces\ and let \[ T: (X, \norm{X}{\cdot}) \rightarrow (Y, \norm{Y}{\cdot}) \] be a linear mapping.
            Then we call $T$ a \highlight{linear operator}.
            We call $T$ \highlight{bounded}, if \[\exists C > 0, \forall x \in X: \norm{Y}{Tx} \le C \norm{X}{x}\]

            \noindent For brevity, we will use the notation \[ T: X \rightarrow Y \] 
            for linear operators rather than \[ T: (X, \norm{X}{\cdot}) \rightarrow (Y, \norm{Y}{\cdot}) \]

        \item
            Let $\normedspace{X}$ be a normed \thefield-vector.
            The (topological) dual space is defined as \[ X' \defeq \dualspace{X} \]

    \end{romanenum}
\end{definition}

\begin{revision} \label{revision:general-revision}
    The following statements are foundational for this topic: \\
    Let \twonormedvectorspaces.
    \begin{romanenum}
        \item The set of continuous, linear operators \contoperators{X}{Y} is a Banach space if and only if $Y$ is a Banach space.
            In particular, the topological dual space \dualspace{X} is a Banach space.
        \item Let $T: X \to Y$ be a linear operator.
            Then $T$ is bounded if and only if $T$ is continous.
        \item Let $(Z, \norm{Z}{\cdot})$ be a normed \thefield-vector space,
            let $T: X \to Y$ be a linear, bounded operator and 
            let $S: Y \to Z$ be a linear, bounded operator.
            Then $S \chain T$ is a linear, bounded operator.
        \item Let $(Z, \norm{Z}{\cdot})$ be a normed \thefield-vector space,
            let $T: X \to Y$ be a linear operator and 
            let $S: Y \to Z$ be a linear operator.
            If $T$ or $S$ is compact, $S \chain T$ is compact.
    \end{romanenum}
\end{revision}
\begin{proof}
    \refertolecture
\end{proof}

\begin{definition}
    Let \twonormedvectorspaces\ and $T \in \contoperators{X}{Y}$. \\
	Then $T': Y' \rightarrow X', y' \mapsto y' \chain T$ is called the adjoint operator.
    From now on, we will implicitly refer to the normed \thefield\ vector spaces $X, Y$ and the topological dual spaces $X', Y'$ when talking about the dual operator $T'$.
\end{definition}

\begin{remark}
    In comparison to the operators we have previously worked with, the adjoint operator takes and outputs linear, bounded operators.
	It's one more level of abstraction removed from \thefield. \\
	For $y' \in Y' = \dualspace{Y}$, the adjoint operator evaluates to $T' y' \in \dualspace{X} = X'$.
\end{remark}

\begin{example}
    % TODO: Add simple integral operator example.
\end{example}